### 2021-10-13 plos_latex_template.tex
Rocío: edits to communication rule

In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. 
% Rocío: we can add our Nature Correspondence letter to the citations later on. 
Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020}.
% Rocío: we can add our Nature Correspondence letter to the citations here too. I would've liked one that talks about increase in participation of tutors and presenters but haven't found any. 
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location. 
% Rocío: main idea of the paragraph: online conferences are more inclusive
% Liz: but only if you make them that way. The R conference was a standout and innovative compared to others I've considred attending since Covid.

Networking and socializing in online conferences may be challenging, mostly because we are used to personal interactions at coffee or lunch breaks, or other in-person social settings \cite{salibaGettingGripsOnline2020}. 
However, social events in in-person conferences can be exclusionary (e.g. galas or dinner nights at expensive venues), and they are a common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. Organizers should invest time in creating opportunities to meet and bond virtually. 
An online event is a great opportunity to create spaces for people who do not enjoy in-person networking. 
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people's limits, preferences, and remember that 'the usual' does not necessarily work for everyone in your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.
Moreover, virtual communication may make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
% Rocío: Networking paragraph.

In addition to being more inclusive, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}. 
Besides, if conference materials are systematically stored and pre-recorded, they can be made available to people who missed them live or even after the conference is over. 
% Rocío: Short paragraph of additional advantages of online conferences that are not about inclusion.

Alternatively, a conference could have a hybrid format with an in-person and an online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%ast: I added Liz's suggestion and rephrased slightly to convey that online is not automatically inclusive
For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Hybrid conferences paragraph.

@@

\section{Express the welcoming spirit in your communication strategy}
\label{rule_communication}
%ast: sara mentioned something that can be good spirit for the communication rule: don't do anything to the outside that you are not doing for your team. it's communication strategy but also there should be coherence between your actions and your words. %Rocío: I love that. Please add it.
An inclusive communication is to give each individual within the conference the opportunity to express themselves and a sense of belonging. Embedding inclusion in your communication strategy can only come by design, it should not be a thoughtless default. In the essence of your communication strategy, there should be coherence between actions and words. Always attend to the needs of your team, in other words, don't do anything to the outside that you are not doing for your team.

As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines [*here] and advertising support for attendees, such as fee waivers and financial support. People who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory. 
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too.

Always include explicit descriptions of accessibility and other inclusion practices (\textbf{Rule \ref{rule_accessibility}}). Help people answer the question "Can I attend this conference?" without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.%ast this explicit description seems to belong to the previous paragraph, maybe where I marked [*here]?
Be careful with the language you use in promoting the conference and how you talk about minoritized people, assertively set tone of inclusion. In addition to excluding derogatory or discriminatory language, make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept feedback without being offended.

As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)

>>>>

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
In-person interaction in conferences is priceless, but only for the ones who can afford to attend. 
Barriers such as cost of registration, transport and accommodation, the logistics of long-distance travel, and discriminatory visa applications, are particularly true for conferences that usually take place in high-income countries \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019,jooKeepOnlineOption2021}. 
Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities \cite{salibaGettingGripsOnline2020}.
This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021, roosOnlineConferencesNew2020, jooKeepOnlineOption2021}.
The online format may also make it easier to be inclusive of geographic regions by encompassing several timezones in the week or easily deciding to rotate the favored timezone year to year without depending on a conference central location.

% Rocío: main idea of the paragraph: In the context of reopening, hybrid conferences may be a thing. They also need to be inclusive.
Within the context of the COVID-19 pandemic, many conferences embraced the online format; but at the time of writing, some are reverting to in-person, which risks going back to the barriers mentioned above.
Alternatively, a conference could have a hybrid format with an in-person and an online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. 
The challenge and requirement for this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.
%Needless to say, an online component is not inclusive \textit{per se} without taking many other precautions, and it does not eliminate the need to organize an inclusive in-person component. 
%For instance, to avoid favoring the same regions in every edition, we suggest rotating the location of the conference—if it is global, moving from one continent to another. 
% Rocío: Now I'm not sure if we need the last two sentences that I commented.

% Rocío: main idea: Networking may seem the weakest part of online conference, but it doesn't have to be. 
% Rocío: I'm still doubting if this should be the 2nd paragraph, and I think that it needs some editing.
Networking and socializing have been mentioned as challenging aspects of online conferences, mostly because we are used to interactions at in-person social settings such as coffee or lunch breaks \cite{salibaGettingGripsOnline2020, roosOnlineConferencesNew2020}. 
However, in-person settings might not be comfortable or appealing to everyone for socializing; 
some of these spaces may be exclusionary (e.g. galas or dinner nights at expensive venues), and common place for code of conduct violations to occur \cite{auroraHowRespondCode2019}. 
On the other hand, there is evidence that virtual communication before face-to-face interaction can make people from minoritized groups feel more included, thus participate more (e.g. \cite{trianaDoesOrderFacetoFace2012}).
Organizers of online conferences should invest time in creating opportunities to meet and bond virtually, respecting people's limits, preferences, and remembering that `the usual' does not necessarily work for everyone, and that one single networking activity will never serve the whole community. 
It is worth trying new and varied activities that might work with subgroups of attendees.
Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with teamwork like trivia, offer some events that can be enjoyed passively like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session.

@@

\section{Express the welcoming spirit in your communication strategy}
\label{rule_communication}

%Rocío: Main idea: Include people in your communication 
Your communication strategy should reflect the spirit of the conference and be inclusive by design. 
As you design the promotion plan, actively reach out and promote the conference to people who have been systematically excluded. 
If part of the community has been historically discriminated against, emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. This could be done through publishing the diversity statement (e.g., https://user2021.r-project.org/about/diversitystatement/), the code of conduct, and accessibility guidelines and advertising financial support, or other creative ways that show that everyone is seen, respected, and welcome (Fig. MARMOTS).
Also, try to advertise the conference in multiple languages (\textbf{Rule \ref{rule_language}}) and multiple platforms (e.g. Twitter, Facebook, LinkedIn, conference website, email lists)--
part of your community may use a different platform than the one who regularly attend the conference. 
With proactive inclusive communication strategies, potential attendees will be assured that this is their space and community too (e.g. having accessibility guidelines on your website sends a welcoming message to people with disabilities), without the emotional labor of contacting the organizers and be public about your commitment to equity and inclusion.

%Rocío: Main idea: Use inclusive language
Inclusive language--language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups--should be used in all communications \cite{hallDesigningDiversityInclusion2019}.
Make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. 
Do not expect minoritized people to teach you--it's not their role--and accept feedback without being offended.

% Rocío: not sure if this belongs to the rule—Andrea make a convincing case :-P
As we mentioned previously, conference attendees should have ample opportunities to network with each other. However, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-led by community leaders, these sessions can be more productive successful, and inclusive than trying to organize one single activity that can please the whole community (\textbf{Rule \ref{rule_unbias}}). Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.
%Adithi had said: "Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference." but i see batool included newbies here :)
